LME-MPM applied to quasi-brittle fracture.

### 2020-05-21 2020_EFM_MPM_Eigensoftening.tex
Cambios 20-5-2020 (I) : Hasta 3.2
Faltan las conclusiones

\documentclass[preprint,12pt,a4paper]{elsarticle}
\usepackage{lineno}
\usepackage{hyperref}
\usepackage{float}
\usepackage{graphicx} % figuras
% \usepackage{subfig}
\usepackage{subfigure} % subfiguras
\usepackage{color}
\usepackage{soul}
\usepackage{cancel}

@@

@@

@@

This problem here presented is devoted to compare how much \acrshort{lme}
approximants are able to improve the result \textit{versus} standard
linear interpolation. It consists in a square plate of size $H = 1$ containing
an initial edge crack of length $0.25 \cdot H$ loaded in a pure mode I
by displacement control on the outer flanks of the plate, Figure
\ref{fig:geometry-cracked-panel-mode-I}.
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{figure}
  \centering
  \includegraphics[width=0.4\textwidth]{Figure-Mode_I}
  \caption{Geometry and boundary condition of the drop-weight impact test.}
  \label{fig:geometry-cracked-panel-mode-I}
\end{figure}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
The constitutive model consider for numerical experiment is a
linear-elastic Hookean material, where Young's modulus $E = 1.06$, the
Poisson's ratio $\mu = 0.333$, and critical energy-release rate $G_c =
0.0001$.
This set of parameters was taken from \cite{Pandolfi_2012}. In
\acrshort{mpm}, two different discretization are required. In one hand
a cartesian grid of nodes is consider with a nodal spacing value of
0.025. On the other hand, the plate will be modeled with a initial
layout of four particles per element and occupying the Gauss quadrature
positions.

Figure \ref{fig:Reactions-cracked-panel-mode-I} clearly shows the
presence of wiggles in both loading and fracture stages
in the reaction-displacements curve when linear interpolation technique is employed. In
contrast, \acrshort{lme} simulation does not exhibit these spurious
oscillations and remains linear. Besides that, linear interpolation
produces stiffer results than \acrshort{lme}. This can be attributed
to imprecision in the stress field due to grid-crossing phenomena.
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{figure}
  \centering
  \includegraphics[width=0.8\textwidth]{Figure-Reactions-Mode-I}
  \caption{Evolution of the reaction forces plotted \textit{versus}
    the imposed displacement for linear interpolation technique and
    the \acrshort{lme} approximants.}
  \label{fig:Reactions-cracked-panel-mode-I}
\end{figure}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
On the other hand, Figure \ref{fig:Stress-cracked-panel-mode-I}
exhibits more clearly the difference between linear interpolation and
\acrshort{lme}, since numerical instability becomes rather significant
when particles crosses the boundary of the element in the linear
case. Although this behavior does not affect significantly to the peak
stress value in a hookean material \cite{Zhang_EE_2020}, if
hyperelastic materials are consider for future simulations,
inaccuracies in the stress field could affect dramatically to the
final result. On the other hand \acrshort{lme} produces soft stress
field evolution even after breaking    
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{figure}
\centering
\subfigure[t = 25]{\includegraphics[width=0.7\textwidth]{./Figure-Stress-0100-mode-I.png}}
\subfigure[t = 60]{\includegraphics[width=0.7\textwidth]{./Figure-Stress-0250-mode-I.png}}
\subfigure[t = 100]{\includegraphics[width=0.7\textwidth]{./Figure-Stress-0398-mode-I.png}}
\caption{Evolution of the stress tensor magnitude for linear
  interpolation (pictures in the left side), and \acrshort{lme}
  (pictures in the right side). Both simulations performed with an
  eigenerosion algorithm.}
\label{fig:Stress-cracked-panel-mode-I}
\end{figure}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
An important consideration regarding the presence of oscillations once
both parts of the panel are separated, Figure
\ref{fig:Reactions-cracked-panel-mode-I}. These phenomena should not be
attributed to the eigenerosion algorithm since the fracture process is
over. They are due to the dynamic nature of the solver.

@@

Once the accuracy of the \acrshort{lme} approximants have been proven
in Section \ref{sec:3.1}, henceforth will be adopted without comparing
with linear interpolation. Therefore this section is devoted

\begin{figure}
  \centering
  \includegraphics[width=0.8\textwidth]{./Figure-Drop-weight}
  \caption{Geometry and boundary condition of the drop-weight impact test.}
  \label{fig:geometry-drop-weight-impact-test}
\end{figure}

@@

\clearpage

\section{Explicit Predictor-Corrector algorithm}
\label{sec:expl-pred-corr}

\begin{algorithm}
  \DontPrintSemicolon
    %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%% º
    \textbf{Update mass matrix :} $ \tens{m}_{I} = N_{Ip}^{k}\ m_p$ \;
    %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%% 
    \textbf{Explicit Newmark Predictor :}
    \begin{equation*}
      \vec{v}_I^{pred} = \frac{ N_{Ip}^{k} m_p (\vec{v}_p^k + (1 - \gamma)\ \Delta t\ \vec{a}_p^k)}{m_I}
    \end{equation*}\;
    %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%% 
    \textbf{Impose essential boundary conditions :} At the fixed
    boundary, set $\vec{v}_{I}^{pred} = 0$.\; 
    %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%% 
    \textbf{Deformation tensor increment calculation :}
    \begin{equation*}
      \Delta \tens{\varepsilon}_{p}^{k+1} = \Delta t\
        \dot{\tens{\varepsilon}_{p}}^{k+1} = \Delta t\ \left[ \vec{v}_{I}^{pred} \otimes
        \Grad{N_{Ip}^{k+1}} \right]^s
    \end{equation*} \;
    %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%% 
    \textbf{Update the density field :} $\rho_p^{k+1} =
    \frac{\rho_p^k}{1 + \mathit{tra}\left[\Delta\tens{\varepsilon}_{p}^{k+1}\right]}.$\;
    %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%% 
    \textbf{Compute stress field and update damage parameter}\;
    %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%% 
    \textbf{Balance of forces calculation :} Calculate the total grid
    nodal force $\vec{f}_{I}^{k+1} = \vec{f}_{I}^{int,k+1} + \vec{f}_{I}^{ext,k+1}$.\;
    %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%% 
    \textbf{Explicit Newmark Corrector :}
    \begin{equation*}
      \vec{v}_{I}^{k+1} = \vec{v}_{I}^{pred} + \gamma\ \Delta t\ \frac{\vec{f}_{I}^{k+1}}{\tens{m}_I^{k+1}}  
    \end{equation*}\;
    %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
    \textbf{Update particles lagrangian quantities :}
    \begin{align*}
      &\vec{a}_p^{k+1} = \frac{N_{Ip}^k\vec{f}_{I}^{k}}{\tens{m}_I^k}\\
      &\vec{v}_p^{k+1} = \vec{v}_p^n + \Delta t\
        \frac{N_{Ip}^k\
        \vec{f}_{I}^{k}}{\tens{m}_I^k}\\
      &\vec{x}_p^{k+1} = \vec{x}_p^n + \Delta t\
         N_{Ip}^k\ \vec{v}_{I}^{k} +
        \frac{1}{2}\Delta t^2\ \frac{N_{Ip}^k\
        \vec{f}_{I}^{k}}{\tens{m}_I^k}
    \end{align*}\;
    %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%% 
    \textbf{Reset nodal values}\;
    %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%% 
    \label{alg-epc}
    \caption{Explicit Predictor-Corrector scheme}
\end{algorithm}

@@

\begin{algorithm}
  \DontPrintSemicolon
  \KwInput{For each $p$, $\epsilon$-neighbourhood, $f_{t,p}$, $h_{\epsilon,p}$, $w_c$}
  \KwOutput{Return damage parameter $\chi := \{\chi_p\}$}
  % \KwData{Testing set $x$}
  $\chi_p \leftarrow \chi_p^{k}$ \;
  \For{$p$ to $N_p$}
  {
    \If{$\chi_p = 0$ \AND $\epsilon_{f,p} = 0$}
    {
      \For{$q \in B_{\epsilon,p}$}
      {
        $\sigma_{q,I} \leftarrow \text{getEigenvaluesOf}(\sigma_{q})$ \;
        \If{$\chi_q < 1$}
        {
          $\sum m_p\sigma_{p,I} \leftarrow \sum m_p\sigma_{p,I} + m_q\sigma_{q,I}$ \;
        }
        $m_p \leftarrow m_p + m_q$ \;
      }
      $\sigma_{p,\epsilon} \leftarrow \frac{1}{m_p} \sum m_p\sigma_{p,I}$\;
      \If{$\sigma_{p,\epsilon} > f_{t,p}$}
      {
        $\epsilon_{f,p} = \epsilon_{I,p}$ \;  
      }
      \ElseIf{$\chi_p \neq 1$ \AND $\epsilon_{f,p} > 0$}
      {
        $\chi_p^{k+1} \leftarrow \min\Big \{1 , \max \{\chi_p^{k},
        \frac{(\epsilon_{I,p}- \epsilon_{f,p})\ h_{\epsilon,p}}{w_c}
        \} \Big \}$ \;
      }
    }
  }
  \caption{Compute damage parameter $\chi_p^{k+1}$}
  \label{alg-eigens}
\end{algorithm}

>>>>

\documentclass[preprint,12pt,a4paper]{elsarticle}
\usepackage{lineno}
\usepackage{hyperref}
\usepackage{float}
\usepackage{graphicx} % figuras
% \usepackage{subfig}
\usepackage{subfigure} % subfiguras
\usepackage{multirow} 
\usepackage{color}
\usepackage{soul}
\usepackage{cancel}

@@

\newglossaryentry{fc}
{
  name=f$_c$,
  description={Compressive strenght}
}

\newglossaryentry{ft}
{
  name=f$_t$,
  description={Tensile strenght}
}

\newglossaryentry{Gf}
{
  name=G$_F$,
  description={Specific fracture energy}
}

\newglossaryentry{E}
{
  name=E,
  description={Elastic modulus}
}

\newglossaryentry{nu}
{
  name=$\nu$,
  description={Poisson ratio}
}

@@

Implementation details can be consulted in \ref{sec:eigens-algor-1}

@@

This problem here presented is devoted to compare how much \acrshort{lme}
approximants are able to improve the result \textit{versus} standard
linear interpolation. It consists in a square plate of size $H = 1$ containing
an initial edge crack of length $0.25 \cdot H$ loaded in a pure mode I
by displacement control on the outer flanks of the plate, Figure
\ref{fig:geometry-cracked-panel-mode-I}.
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{figure}
  \centering
  \includegraphics[width=0.4\textwidth]{Figure-Mode_I}
  \caption{Geometry and boundary condition of the drop-weight impact test.}
  \label{fig:geometry-cracked-panel-mode-I}
\end{figure}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
The constitutive model consider for numerical experiment is a
linear-elastic Hookean material, where Young's modulus \gls{E} = 1.06, the
Poisson's ratio \gls{nu} = 0.333, and critical energy-release rate
\gls{Gf} = 0.0001. This set of parameters was taken from \cite{Pandolfi_2012}. In
\acrshort{mpm}, two different discretization are required. In one hand
a cartesian grid of nodes is consider with a nodal spacing value of
0.025. On the other hand, the plate will be modeled with a initial
layout of four particles per element and occupying the Gauss quadrature
positions.\\

Figure \ref{fig:Reactions-cracked-panel-mode-I} clearly shows the
presence of wiggles in both loading and fracture stages
in the reaction-displacements curve when linear interpolation technique is employed. In
contrast, \acrshort{lme} simulation does not exhibit these spurious
oscillations and remains linear. Besides that, linear interpolation
produces stiffer results than \acrshort{lme}. This can be attributed
to imprecision in the stress field due to grid-crossing phenomena.\\
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{figure}
  \centering
  \includegraphics[width=0.8\textwidth]{Figure-Reactions-Mode-I}
  \caption{Evolution of the reaction forces plotted \textit{versus}
    the imposed displacement for linear interpolation technique and
    the \acrshort{lme} approximants.}
  \label{fig:Reactions-cracked-panel-mode-I}
\end{figure}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
On the other hand, Figure \ref{fig:Stress-cracked-panel-mode-I}
exhibits more clearly the difference between linear interpolation and
\acrshort{lme}, since numerical instability becomes rather significant
when particles crosses the boundary of the element in the linear
case. Although this behavior does not affect significantly to the peak
stress value in a hookean material \cite{Zhang_EE_2020}, if
hyperelastic materials are consider for future simulations,
inaccuracies in the stress field could affect dramatically to the
final result. On the other hand \acrshort{lme} produces soft stress
field evolution even after breaking.
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{figure}
\centering
\subfigure[t = 25]{\includegraphics[width=0.73\textwidth]{./Figure-Stress-0100-mode-I.png}}
\subfigure[t = 60]{\includegraphics[width=0.73\textwidth]{./Figure-Stress-0250-mode-I.png}}
\subfigure[t = 100]{\includegraphics[width=0.73\textwidth]{./Figure-Stress-0398-mode-I.png}}
\caption{Evolution of the stress tensor magnitude for linear
  interpolation (pictures in the left side), and \acrshort{lme}
  (pictures in the right side). Both simulations performed with an
  eigenerosion algorithm.}
\label{fig:Stress-cracked-panel-mode-I}
\end{figure}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
An important consideration regarding the presence of oscillations once
both parts of the panel are separated, Figure
\ref{fig:Reactions-cracked-panel-mode-I}. These phenomena should not be
attributed to the eigenerosion algorithm since the fracture process is
over. They are due to the dynamic nature of the solver.

@@

Once the accuracy of the \acrshort{lme} approximants have been proven
in Section \ref{sec:3.1}, henceforth will be adopted without comparing
with linear interpolation. Therefore this section is devoted to proof
the accuracy of the eigensoftening algorithm to simulate the behavior
of quasi-brittle materials under dynamic loading. An example of this
load case is the three-point bending test on a notched beam is
conducted under impact loading. This kinds of experiments are
challenging to reproduce numerically with an explicit solver since
fracture occurs in a period of milliseconds and high strength
materials are involved. Accordingly, really small time steps are
required to achieve a simulation numerically stable. Additionally,
this kind of simulations traditionally require the implementation of
contact algorithm to reproduce the hammer impact in the beam. In the
\acrshort{mpm}, the implementation of this feature can be bypassed to
obtain preliminary results. Nevertheless, this kind of algorithm could
be implemented to get ever more accurate solutions. Although this is
out of the scope of this document, future research of authors will overcome this
limitation.\\

As in the study of Zhang {\it et al.} \cite{Zhang_2009,Zhang_2010a},
an impact hammer of 120.6 kg was employed to drop it at an impact
speed of 881 mm/s. The beam dimensions where 100 mm x 100 mm (B x D)
in cross section, and 420 mm in total lenght (L). The initial
notch-depth ratio was approximately 0.5, and the span, S, was fixed at
300 mm during the tests, see Figure
\ref{fig:geometry-drop-weight-impact-test}.
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{figure}
  \centering
  \includegraphics[width=0.8\textwidth]{./Figure-impact-test}
  \caption{Geometry and boundary condition of the drop-weight impact test.}
  \label{fig:geometry-drop-weight-impact-test}
\end{figure}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
The material adopted for the simulation was characterised by Navas
{\it et al.}~\cite{Navas_2017_ES}, and the material properties, such
as the material density, \gls{rho}, the compressive strenght, \gls{fc},
the tensile strenght \gls{ft}, the specific fracture energy, \gls{Gf},
and the elastic modulus \gls{E} are given in table \ref{tab:concrete-properties}.\\
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\begin{table}
  \centering
  \begin{tabular}[]{c c c c c c c}
    \hline
      &   $\rho$   & $\text{f}_c$ & $\text{f}_t$ & $G_F$ &   E   & $\text{d}_{max}$ \\
      & (kg/m$^3$) &     (MPa)    &     (MPa)    & (N/m) & (GPa) & (mm) \\
    \hline
Value &    2368    &     102.7    &      5.4     &  141  &   31  &  12 \\
    \hline
  \end{tabular}
  \caption[Mechanical properties of thje concrete]{Mechanical
    properties of the high strenght concrete.}
  \label{tab:concrete-properties}
\end{table}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
For the present research, a 2D setting of the experiment has been
adopted. Both the hammer and the concrete beam are explicitly
represented. Several levels of discretization are employed to assess
the objectiveness of the obtained results. The results here exposed
are from a discretization of 30616 material points and 6297 nodes. To
achieve better results, an unestructured grid layout has been adopted
focusing the minimum nodal spacing, 0.47 mm, in the middle of the beam and in
the impact surface of the hammer. For this simulation, $\gamma$ was
fixed to 6.\\

First the reaction and impact forces are validated against their
experimental counterparts. Since the impact forces applied by the
hammer and the reaction forces at the two supports were experimentally
measured, they are compared with numerical ones in Figure
\ref{fig:Reactions-Forces-impact-test}. 
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%% 
\begin{figure}
  \centering
  \includegraphics[width=0.8\textwidth]{Figure-impact-test-Forces-Time}
  \caption{Evolution of the reaction forces plotted \textit{versus}
    the imposed displacement for linear interpolation technique and
    the \acrshort{lme} approximants.}
  \label{fig:Reactions-Forces-impact-test}
\end{figure}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
Note that the general trend of both forces are correctly
captured. Observe that in experimental solution the delay between
impact peak and reactions peak does not agree with the numerical
ones. On the other hand, if the delay is computed analytically through
concrete wave speed is much closer to the numerical results. A
feasible explanation to this unexpected result can due the composite
nature of the concrete. Another discrepancy is in the much higher
values of maximum impact and reaction forces in numerical than in
experimental. In this case, this is a consequence of over-rigidity
supports. Therefore, this results can be improved in future research
by a suitable contact algorithm.\\

And finally horizontal stress distributions and crack propagation are
extracted, see Figure \ref{fig:Stress-vs-damage-impact-test}. Compared
with results obtained by \cite{Navas_2017_ES}, stress evolution

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{figure}
\centering
\subfigure[t = 0.27 ms]{\includegraphics[width=0.83\textwidth]{./Figure-impact-test-Damage-vs-Stress-0060}}
\subfigure[t = 0.36 ms]{\includegraphics[width=0.83\textwidth]{./Figure-impact-test-Damage-vs-Stress-0080}}
\subfigure[t = 0.45 ms]{\includegraphics[width=0.83\textwidth]{./Figure-impact-test-Damage-vs-Stress-0100}}
\caption{Evolution of the stress tensor magnitude for linear
  interpolation (pictures in the left side), and \acrshort{lme}
  (pictures in the right side). Both simulations performed with an
  eigenerosion algorithm.}
\label{fig:Stress-vs-damage-impact-test}
\end{figure}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

@@

\section{Explicit Newmark Predictor-Corrector algorithm}
\label{sec:expl-pred-corr}
\begin{algorithm}[H]
  \DontPrintSemicolon
    %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%% º
    \textbf{Update mass matrix :} $ \tens{m}_{I} = N_{Ip}^{k}\ m_p$ \;
    %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%% 
    \textbf{Explicit Newmark Predictor :}
    \begin{equation*}
      \vec{v}_I^{pred} = \frac{ N_{Ip}^{k} m_p (\vec{v}_p^k + (1 - \gamma)\ \Delta t\ \vec{a}_p^k)}{m_I}
    \end{equation*}\;
    %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%% 
    \textbf{Impose essential boundary conditions :} At the fixed
    boundary, set $\vec{v}_{I}^{pred} = 0$.\; 
    %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%% 
    \textbf{Deformation tensor increment calculation :}
    \begin{equation*}
      \Delta \tens{\varepsilon}_{p}^{k+1} = \Delta t\
        \dot{\tens{\varepsilon}_{p}}^{k+1} = \Delta t\ \left[ \vec{v}_{I}^{pred} \otimes
        \Grad{N_{Ip}^{k+1}} \right]^s
    \end{equation*} \;
    %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%% 
    \textbf{Update the density field :} $\rho_p^{k+1} =
    \frac{\rho_p^k}{1 + \mathit{tra}\left[\Delta\tens{\varepsilon}_{p}^{k+1}\right]}.$\;
    %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%% 
    \textbf{Compute stress field and update damage parameter}\;
    %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%% 
    \textbf{Balance of forces calculation :} Calculate the total grid
    nodal force $\vec{f}_{I}^{k+1} = \vec{f}_{I}^{int,k+1} + \vec{f}_{I}^{ext,k+1}$.\;
    %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%% 
    \textbf{Explicit Newmark Corrector :}
    \begin{equation*}
      \vec{v}_{I}^{k+1} = \vec{v}_{I}^{pred} + \gamma\ \Delta t\ \frac{\vec{f}_{I}^{k+1}}{\tens{m}_I^{k+1}}  
    \end{equation*}\;
    %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
    \textbf{Update particles lagrangian quantities :}
    \begin{align*}
      &\vec{a}_p^{k+1} = \frac{N_{Ip}^k\vec{f}_{I}^{k}}{\tens{m}_I^k}\\
      &\vec{v}_p^{k+1} = \vec{v}_p^n + \Delta t\
        \frac{N_{Ip}^k\
        \vec{f}_{I}^{k}}{\tens{m}_I^k}\\
      &\vec{x}_p^{k+1} = \vec{x}_p^n + \Delta t\
         N_{Ip}^k\ \vec{v}_{I}^{k} +
        \frac{1}{2}\Delta t^2\ \frac{N_{Ip}^k\
        \vec{f}_{I}^{k}}{\tens{m}_I^k}
    \end{align*}\;
    %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%% 
    \textbf{Reset nodal values}\;
    %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%% 
    \label{alg-epc}
    \caption{Explicit Newmark Predictor-Corrector scheme}
\end{algorithm}

@@

\begin{algorithm}[H]
  \DontPrintSemicolon
  \KwInput{For each $p$, $\epsilon$-neighbourhood, $f_{t,p}$, $h_{\epsilon,p}$, $w_c$}
  \KwOutput{Return damage parameter $\chi := \{\chi_p\}$}
  % \KwData{Testing set $x$}
  $\chi_p \leftarrow \chi_p^{k}$ \;
  \For{$p$ to $N_p$}
  {
    \If{$\chi_p = 0$ \AND $\epsilon_{f,p} = 0$}
    {
      \For{$q \in B_{\epsilon,p}$}
      {
        $\sigma_{q,I} \leftarrow \text{getEigenvaluesOf}(\sigma_{q})$ \;
        \If{$\chi_q < 1$}
        {
          $\sum m_p\sigma_{p,I} \leftarrow \sum m_p\sigma_{p,I} + m_q\sigma_{q,I}$ \;
        }
        $m_p \leftarrow m_p + m_q$ \;
      }
      $\sigma_{p,\epsilon} \leftarrow \frac{1}{m_p} \sum m_p\sigma_{p,I}$\;
      \If{$\sigma_{p,\epsilon} > f_{t,p}$}
      {
        $\epsilon_{f,p} = \epsilon_{I,p}$ \;  
      }
      \ElseIf{$\chi_p \neq 1$ \AND $\epsilon_{f,p} > 0$}
      {
        $\chi_p^{k+1} \leftarrow \min\Big \{1 , \max \{\chi_p^{k},
        \frac{(\epsilon_{I,p}- \epsilon_{f,p})\ h_{\epsilon,p}}{w_c}
        \} \Big \}$ \;
      }
    }
  }
  \caption{Compute damage parameter $\chi_p^{k+1}$}
  \label{alg-eigens}
\end{algorithm}
